---
title: "Untitled"
author: "Anaya Hall"
date: "4/19/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**
# -FUNCTIONS-

##### White's Test
Regress the squared residuals ($e^2_i$) on a constant, all variables in *$X$*, squares of all variables in *$X$* and all cross products. $n \dot R^2$ from this regression is distributed as a $\chi^2_{(p-1)}$, where p is the number of regressors in this equation including the constant. The null in this test is homoskedastic disturbances.


``` {r white_test_fxn, include = F}

white_test <- function(resid, cov_mat){
  
  cov_mat %<>% as.matrix()
  
  # Interaction matrix
  cov_n <- nrow(cov_mat)
  cov_k <- sum(seq(ncol(cov_mat)))
  int_mat <- matrix(NA, nrow = cov_n, ncol = cov_k)
  
  # Loop through all columns to create interaction matrix
  for (i in 1:ncol(cov_mat)) {
    for (j in i:ncol(cov_mat)) {
      if (i == 1) m <- j
      if (i > 1) m <- sum(seq(ncol(cov_mat), 1, -1)[1:(i-1)]) + (j - i + 1)
      int_mat[, m] <- cov_mat[, i] * cov_mat[, j]
    }
  }
  
  # Bind together with covariate matrix
  cov_mat %<>% cbind(.,int_mat)
  # Make sure unique (see documentation for MARGIN = 2)
  cov_mat %<>% unique(MARGIN = 2)
  # Add intercept (column of ones)
  cov_mat %<>% cbind(1,.)

  # Outcome var ('y') is squared residual
  y_data <- resid^2
  
  # y-hat for residual regression = X*beta  
  y_hat <- cov_mat %*% solve(t(cov_mat) %*% cov_mat) %*% t(cov_mat) %*% y_data 
  
  # Calculate SSM and SST for R^2
  SSM <- sum((y_hat - mean(y_data))^2)
  SST <- sum((y_data - mean(y_data))^2)
  
  # Calculate White test statistic = R^2 * n
  test_stat <- SSM / SST * cov_n
  # Calculate pvalue
  pvalue <- 1 - pchisq(test_stat, df = (ncol(cov_mat)-1)) #dof is p-1 yes?
  
  
  return(list(PValue = pvalue, TestStat = test_stat, dof=ncol(cov_mat)))
  
  # white test results
  # whitetest_df <- data.frame()
  # 
  # 
  # white_table <-  whitetest_df %>% knitr::kable(
  #     booktabs = T,
  #     format.args = list(scientific = F),
  #     escape = F,
  #     caption = "White Test")
  
}


```



##### Goldfeld-Quant Test:
``` {r goldfeldquant_fxn}

GQ_test <- function(e1, e2, k) {
  n1 <- length(e1)
  n2 <- length(e2)
  
  SSE1 <- (t(e1) %*% e1)/(n1-k)
  SSE2 <- (t(e2) %*% e2)/(n2-k)
  test_stat <- SSE1 / SSE2
  
  pvalue <- 1 - pf(test_stat, n1-k, n2-k)
  
  return(data.frame(
    "Test Statistic" = test_stat,
    "P-Value" = pvalue
  ))
}

```

```{r 2c, include = FALSE}
# Prep for test
# Rank by tenure
wage_df %<>% arrange(tenure)
# Splitting the data, removing the middle 235 observations
wage_df1 <- wage_df[1:350,]
wage_df2 <- wage_df[586:935,]
# Run two regressions saving their residuals
gq_resid_1 <- ols(wage_df1, y_data = "lwage", 
               X_data = c("exper", "tenure", "married", "south", "urban", "black", "educ"))$vars$e
gq_resid_2 <- ols(wage_df2, y_data = "lwage", 
               X_data = c("exper", "tenure", "married", "south", "urban", "black", "educ"))$vars$e

# Run GQ Test
GQ_test(gq_resid_1, gq_resid_2, k=7) %>% knitr::kable()

```

##### Breusch-Pagan Test for heteroskedastic errors
```{r breushpagan_fxn}
BP_test <- function(data, e, cov_mat) {
  
  df <- data
  
  n <- length(e)
  
  dep_var <- (n * e^2 / sum(e^2))
  
  bp_df <- data.frame(dep_var, cov_mat)
  
  bp_Xs <- names(bp_df[2:ncol(bp_df)])
  # run ols & save residuals
  bp_ols <- ols(bp_df, y = "dep_var", X = bp_Xs)
  yhat <- bp_ols$vars$y_hat
  
  ESS <- sum((dep_var - yhat)^2)
  
  test_stat <- (1/2)*ESS
  
  pvalue <- 1 - pchisq(test_stat, df = nrow(cov_mat))
  
  return(data.frame(
    "Test Statistic" = test_stat,
    "P-Value" = pvalue
  ))
}

```


##### White Robust Standard Errors
```{r Spherical varcov fxn}
# Function for OLS coefficient estimates
b_ols <- function(y, X) {
  # Calculate beta hat
  beta_hat <- solve(t(X) %*% X) %*% t(X) %*% y
  # Return beta_hat
  return(beta_hat)
}

# Function for OLS coef., SE, t-stat, and p-value
vcov_ols <- function(data, y_var, X_vars, intercept = T) {
  # Turn data into matrices
  y <- to_matrix(data, y_var)
  X <- to_matrix(data, X_vars)
  # Add intercept
  if (intercept == T) X <- cbind(1, X)
  # Calculate n and k for degrees of freedom
  n <- nrow(X)
  k <- ncol(X)
  # Estimate coefficients
  b <- b_ols(y, X)
  # Update names
  if (intercept == T) rownames(b)[1] <- "Intercept"
  # Calculate OLS residuals
  e <- y - X %*% b
  # Calculate s^2
  s2 <- (t(e) %*% e) / (n-k)
  s2 %<>% as.numeric()
  # Inverse of X'X
  XX_inv <- solve(t(X) %*% X)
  # Return the results
  return(as.numeric(s2) * XX_inv)
}  
```

```{r Robust varcov fxn}
vcov_white <- function(data, y_var, X_vars, intercept = T) {
  # Turn data into matrices
  y <- to_matrix(data, y_var)
  X <- to_matrix(data, X_vars)
  # Add intercept
  if (intercept == T) X <- cbind(1, X)
  # Calculate n and k for degrees of freedom
  n <- nrow(X)
  k <- ncol(X)
  # Estimate coefficients
  b <- b_ols(y, X)
  # Update names
  if (intercept == T) rownames(b)[1] <- "Intercept"
  # Calculate OLS residuals
  e <- y - X %*% b
  # Inverse of X'X
  XX_inv <- solve(t(X) %*% X)
  # For each row, calculate x_i' x_i e_i^2; then sum
  sigma_hat <- lapply(X = 1:n, FUN = function(i) {
    # Define x_i
    x_i <- matrix(as.vector(X[i,]), nrow = 1)
    # Return x_i' x_i e_i^2
    return(t(x_i) %*% x_i * e[i]^2)
  }) %>% Reduce(f = "+", x = .)
  # Return the results
  return(XX_inv %*% sigma_hat %*% XX_inv)
}
```


##### FGLS
```{r fglsfunctionattempt, include=F}
# ATTEMPT TO MAKE A FUNCTION --- do not include!
# fgls <- function(data, y_data, X_data) {
#   
#   y <- to_matrix(data, y_data)
#   X <- to_matrix(data, X_data) %>% cbind(1,.)
#   Z <- to_matrix(data, X_data) %>% cbind(1,.)
#   
#   # run first regression & save resid
#   e <- ols(y, X)$vars$e
#   
#   w <- ols(e^2, Z)$vars$y_hat
#   
#   W <- (1/sqrt(w))
#   
#   X_tilde <- X %*% X
#   y_tilde <- y %*% W
#   
#   fgls_df <- cbind(X_tilde, y_tilde)
#   
#   fgls_results <- ols(fgls_df, colnames(tilde)[1], colnames(tilde)[2:ncol(tilde)], intercept = F)
#   
#   return(fgls_results)
# }
# 
# #run function on data
# 
# fgls(wage_df, wage_df$lwage, cov_mat)
```

